---
title: "Dissertation"
author: "Jed Green"
date: "2024-04-22"
output: word_document
---

# Dissertation markdown script:

This is the R markdown for my dissertation titled: Psychological Borders: Exploring Attitudes Towards Migrants in the UK from a Psychological Perspective. I have made this (in conjunction with making all my code publicly available) with the intent of making my research more transparent and reproducible.

All of the data used was downloaded from the original source and has not been re-uploaded here. To gain access to it please visit: <https://www.sheffield.ac.uk/psychology-consortium-covid19>

### Packages:

Below is a list of all packages used:

```{r}library(haven)}
library(ggplot2)
library(stargazer)
library(ltm)
library(psych)
library(dplyr)
library(car)
library(corrplot)
library(summarytools)
```

### Importing, filtering and merging the data:

Not all of the measures used in this paper were available in the same wave of the survey. This is due to certain metrics only being asked to new respondents in wave 4 (such as educational level and political scale). Therefore, we need to merge this missing data from previous waves when these respondents answered these exact same questions as posited in wave 4. Below is the code we used to do this.

```{r}
# Loading in main W4 Data:
original_df<-read_sav("C19PRC_UK_W4_archive_final.sav")

# Loading in previous waves data with the anwsers to the missing W4 data
edudf<-read_sav("C19PRC_UKW1W2_archive_final.sav")
edudf1<-read_sav("C19PRC_UK_W3_archive_final.sav")

# Filtering dataframes to contain only the variables we wish to merge
ldf1<-as.data.frame(cbind(edudf$pid,edudf$W1_Education, edudf$W1_Political_Scale))
colnames(ldf1)<-c("pid","W1_Edu", "W1_Political_Scale")

ldf2<-as.data.frame(cbind(edudf1$pid,edudf1$W3_Education, edudf1$W3_Political_Scale))
colnames(ldf2)<-c("pid","W2_Edu","W3_Political_Scale")

# Merging all of the data into a single dataframe
mdf1<-merge.data.frame(original_df,ldf1, by ="pid", all = T)
mdf2<-merge.data.frame(mdf1,ldf2, by ="pid", all = T)


# Replace missing values in W4_Education using responses from W1 or W2
mdf2 <- mdf2 %>%
  mutate(W4_Education = ifelse(is.na(W4_Education), ifelse(W4_Type == 1, coalesce(W1_Edu, W2_Edu), NA), W4_Education))

# Merge W1_Edu and W2_Edu into W4_Education based on conditions
mdf2 <- mdf2 %>%
  mutate(W4_Education = ifelse(W4_Type == 1 & is.na(W4_Education), coalesce(W1_Edu, W2_Edu), W4_Education)) # Filter out rows where W4_Type is NA

# Remove the added variables from previous waves which are no longer needed
mdf2 <- mdf2 %>%
  select(-W1_Edu, -W2_Edu)
mdf2$W4_Education <- as.numeric(as.character(mdf2$W4_Education))

# Replace missing values in W4_Political_Scale using responses from W1 or W2
mdf2 <- mdf2 %>%
  mutate(W4_Political_scale = ifelse(is.na(W4_Political_scale), ifelse(W4_Type == 1, coalesce(W1_Political_Scale, W3_Political_Scale), NA), W4_Political_scale))

# Merge W1_Political_Scale and W3_Political_Scale into W4_Political_scale based on conditions
mdf2 <- mdf2 %>%
  mutate(W4_Political_scale = ifelse(W4_Type == 1 & is.na(W4_Political_scale), coalesce(W1_Political_Scale, W3_Political_Scale), W4_Political_scale)) %>%
  filter(!is.na(W4_Type))

# Remove the added variables from previous waves which are no longer needed
mdf2 <- mdf2 %>%
  select(-W1_Political_Scale, -W3_Political_Scale)
mdf2$W4_Political_scale<- as.numeric(as.character(mdf2$W4_Political_scale))

# Rename dataframe for ease of use
df<-mdf2

# Remove anyone who isn't born in the UK
df<- subset(df,df$W4_WhereBorn != 5)
```

### Coding dependent variable:

In order to create our dependent variable we first need to re-scale some of measures and then add them into an additive scale

```{r}
# Table of 1st dependent measure before re-scaling
table(df$W4_MigrantAttitudes1)

# Define the transformation function used for re-scaling
transform_likert <- function(x) {
  transformed_value <- ceiling(x / 2)
  return(transformed_value)}

# Apply the transformation function to the measure
df$W4_MigrantAttitudes1_rescaled <- transform_likert(df$W4_MigrantAttitudes1)

# Print the transformed data to ensure it worked 
table(df$W4_MigrantAttitudes1_rescaled)


# Table of 2nd dependent measure before re-scaling
table(df$W4_MigrantAttitudes2)

# Apply the transformation function to the measure
df$W4_MigrantAttitudes2_rescaled <- transform_likert(df$W4_MigrantAttitudes2)

# Print the transformed data to ensure it worked 
table(df$W4_MigrantAttitudes2_rescaled)

# Create additive Dependent variable

df$W4_TotalMigrantAttitudes <- rowSums(df[, c("W4_MigrantAttitudes1_rescaled", 
                                 "W4_MigrantAttitudes2_rescaled", 
                                 "W4_Immigration_1", 
                                 "W4_Immigration_1")])

summary(df$W4_TotalMigrantAttitudes)

# Calculate cronach alpha

dependent_data<-data.frame(df[, c("W4_MigrantAttitudes1_rescaled", 
                               "W4_MigrantAttitudes2_rescaled", 
                               "W4_Immigration_1", 
                               "W4_Immigration_1")])
cronbach.alpha(dependent_data)
```

### Calculating cronbach alpha for independent variables:

```{r}
Extraversion_data<-data.frame(df[, c("W4_Personality1R",
                                     "W4_Personality6")])
cronbach.alpha(Extraversion_data)

Agreeableness_data<-data.frame(df[, c("W4_Personality2",
                                     "W4_Personality7R")])
cronbach.alpha(Agreeableness_data)

Conscientiousnes_data<-data.frame(df[, c("W4_Personality3R",
                                      "W4_Personality8")])
cronbach.alpha(Conscientiousnes_data)

Neuroticism_data<-data.frame(df[, c("W4_Personality4R",
                                    "W4_Personality9")])
cronbach.alpha(Neuroticism_data)

Openness_data<-data.frame(df[, c("W4_Personality5R",
                                    "W4_Personality10")])
cronbach.alpha(Openness_data)

Depression_data<-data.frame(df[, c("W4_Dep1",
                                   "W4_Dep2",
                                   "W4_Dep3",
                                   "W4_Dep4",
                                   "W4_Dep5",
                                   "W4_Dep6",
                                   "W4_Dep7",
                                   "W4_Dep8",
                                   "W4_Dep9")])
cronbach.alpha(Depression_data)


Paranoia_data<-data.frame(df[, c("W4_Paranoia1",
                                   "W4_Paranoia2",
                                   "W4_Paranoia3",
                                   "W4_Paranoia4",
                                   "W4_Paranoia5")])
cronbach.alpha(Paranoia_data)

Conspiracy_data<-data.frame(df[, c("W4_Conspiracy_1",
                                 "W4_Conspiracy_2",
                                 "W4_Conspiracy_3",
                                 "W4_Conspiracy_4",
                                 "W4_Conspiracy_5")])
cronbach.alpha(Conspiracy_data)
```

### Re-code independent variables:

We need to re-format our variables to make them conducive for our descriptive statistics and OLS regression

```{r}
# Making Gender into a binary
table(df$W4_Gender)
df$W4_Gender_Binary<- NA
df$W4_Gender_Binary[df$W4_Gender == 1] <- 1
df$W4_Gender_Binary[df$W4_Gender == 2] <- 0
table(df$W4_Gender_Binary)

# Making visiable minority ino a binary
table(df$W4_Ethnicity)
df$W4_Visible_minority<- NA
df$W4_Visible_minority<-ifelse(df$W4_Ethnicity == 1 | df$W4_Ethnicity == 2, 0,1)
table(df$W4_Visible_minority)

# Re-coding education into a catagorical variable
table(df$W4_Education)
df$W4_Education_categories<-NA
df$W4_Education_categories[df$W4_Education == 1] <- "No Qualifications"
df$W4_Education_categories[df$W4_Education == 2] <- "O-Level/GCSE or A-Level or Technical qualification"
df$W4_Education_categories[df$W4_Education == 3] <- "O-Level/GCSE or A-Level or Technical qualification"
df$W4_Education_categories[df$W4_Education == 4] <- "O-Level/GCSE or A-Level or Technical qualification"
df$W4_Education_categories[df$W4_Education == 5] <- "Undergraduate degree or Diploma"
df$W4_Education_categories[df$W4_Education == 6] <- "Undergraduate degree or Diploma"
df$W4_Education_categories[df$W4_Education == 7] <- "Postgraduate degree"
df$W4_Education_categories[df$W4_Education == 8] <- "Other"
df$W4_Education_categories<- factor(df$W4_Education_categories, levels = c("No Qualifications", "O-Level/GCSE or A-Level or Technical qualification", "Undergraduate degree or Diploma","Postgraduate degree", "Other"))
table(df$W4_Education_categories)

# Re-coding education into a numerical catagorical variable
table(df$W4_Education)
df$W4_Education_categories_Num<-NA
df$W4_Education_categories_Num[df$W4_Education == 1] <- "1"
df$W4_Education_categories_Num[df$W4_Education == 2] <- "2"
df$W4_Education_categories_Num[df$W4_Education == 3] <- "2"
df$W4_Education_categories_Num[df$W4_Education == 4] <- "2"
df$W4_Education_categories_Num[df$W4_Education == 5] <- "3"
df$W4_Education_categories_Num[df$W4_Education == 6] <- "3"
df$W4_Education_categories_Num[df$W4_Education == 7] <- "4"
df$W4_Education_categories_Num[df$W4_Education == 8] <- "5"
df$W4_Education_categories_Num<- factor(df$W4_Education_categories_Num, levels = c(1, 2, 3, 4, 5))
table(df$W4_Education_categories_Num)

# Re-coding education into a numerical catagorical variable without other group
table(df$W4_Education)
df$W4_Education_categories_Num1<-NA
df$W4_Education_categories_Num1[df$W4_Education == 1] <- "1"
df$W4_Education_categories_Num1[df$W4_Education == 2] <- "2"
df$W4_Education_categories_Num1[df$W4_Education == 3] <- "2"
df$W4_Education_categories_Num1[df$W4_Education == 4] <- "2"
df$W4_Education_categories_Num1[df$W4_Education == 5] <- "3"
df$W4_Education_categories_Num1[df$W4_Education == 6] <- "3"
df$W4_Education_categories_Num1[df$W4_Education == 7] <- "4"
df$W4_Education_categories_Num1<- factor(df$W4_Education_categories_Num1, levels = c(1, 2, 3, 4))
table(df$W4_Education_categories_Num1)

# Re-coding employment into a catagorical variable
table(df$W4_Employment)
df$W4_Employment_categories<-NA
df$W4_Employment_categories[df$W4_Employment == 1] <- "Employed"
df$W4_Employment_categories[df$W4_Employment == 2] <- "Employed"
df$W4_Employment_categories[df$W4_Employment == 3] <- "Employed"
df$W4_Employment_categories[df$W4_Employment == 4] <- "Employed"
df$W4_Employment_categories[df$W4_Employment == 5] <- "Unemployed"
df$W4_Employment_categories[df$W4_Employment == 6] <- "Unemployed"
df$W4_Employment_categories[df$W4_Employment == 7] <- "Unemployed"
df$W4_Employment_categories[df$W4_Employment == 8] <- "Other Situations"
df$W4_Employment_categories[df$W4_Employment == 9] <- "Other Situations"
df$W4_Employment_categories[df$W4_Employment == 10] <- "Student"
df$W4_Employment_categories<- factor(df$W4_Employment_categories, levels = c("Unemployed", "Employed", "Other Situations", "Student"))
table(df$W4_Employment_categories)

# Re-coding employment into a numerical catagorical variable
table(df$W4_Employment)
df$W4_Employment_categories_Num<-NA
df$W4_Employment_categories_Num[df$W4_Employment == 1] <- "1"
df$W4_Employment_categories_Num[df$W4_Employment == 2] <- "1"
df$W4_Employment_categories_Num[df$W4_Employment == 3] <- "1"
df$W4_Employment_categories_Num[df$W4_Employment == 4] <- "1"
df$W4_Employment_categories_Num[df$W4_Employment == 5] <- "2"
df$W4_Employment_categories_Num[df$W4_Employment == 6] <- "2"
df$W4_Employment_categories_Num[df$W4_Employment == 7] <- "2"
df$W4_Employment_categories_Num[df$W4_Employment == 8] <- "2"
df$W4_Employment_categories_Num[df$W4_Employment == 9] <- "3"
df$W4_Employment_categories_Num[df$W4_Employment == 10] <- "4"
df$W4_Employment_categories_Num<- factor(df$W4_Employment_categories_Num, levels = c(2, 1, 3, 4))
table(df$W4_Employment_categories_Num)

## Re-coding Income into a  catagorical variable
table(df$W4_Income_2019)
df$W4_Income_2019_categories<-NA
df$W4_Income_2019_categories[df$W4_Income_2019 == 1] <- "£0-15,490 per year"
df$W4_Income_2019_categories[df$W4_Income_2019 == 2] <- "£15,491-£25,340 per year"
df$W4_Income_2019_categories[df$W4_Income_2019 == 3] <- "£25,341-£38,740 per year"
df$W4_Income_2019_categories[df$W4_Income_2019 == 4] <- "£38,741-£57,930 per year"
df$W4_Income_2019_categories[df$W4_Income_2019 == 5] <- "£57,931 or more per year"
df$W4_Income_2019_categories<- factor(df$W4_Income_2019_categories)
table(df$W4_Income_2019_categories)
```

### Descriptive statistics:

Now we need to create our descriptive statistics, we made a descriptive table and a correlation matrix

```{r}
# Create a df for variables in descriptive statistics table
cor_df<-data.frame(df[, c("W4_Age_year",
                          "W4_Gender_Binary",
                          "W4_Visible_minority",
                          "W4_Education_categories_Num",
                          "W4_Employment_categories_Num",
                          "W4_Income_2019",
                          "W4_Political_scale",
                          "W4_Extraversion_Total",
                          "W4_Agreeable_Total",
                          "W4_Conscientious_Total",
                          "W4_Neuroticism_Total",
                          "W4_Openness_Total",
                          "W4_Dep_Total",
                          "W4_PTSDdx",
                          "W4_Paranoia_Total",
                          "W4_Conspiracy_Total",
                          "W4_TotalMigrantAttitudes")])
# Give them df new lables
new_names <- c("Age (Years)", "Male", "Visible minority", "Education", "Employment", "Income", "Political scale", "Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness", "Depression", "PTSD_dx", "Paranoia", "Conspiracy", "Migrant Attitudes")
names(cor_df) <- new_names


### create Descriptive Stats and export to CSV:
comprehensive_summary <- describe(cor_df, na.rm = T)
write.csv(comprehensive_summary, "comprehensive_summary.csv")

# Create a df for variables in correlation matrix table
cor_df1<-data.frame(df[, c("W4_Age_year",
                          "W4_Gender_Binary",
                          "W4_Visible_minority",
                          "W4_Education_categories_Num1",
                          "W4_Income_2019_1",
                          "W4_Political_scale",
                          "W4_Extraversion_Total",
                          "W4_Agreeable_Total",
                          "W4_Conscientious_Total",
                          "W4_Neuroticism_Total",
                          "W4_Openness_Total",
                          "W4_Dep_Total",
                          "W4_PTSDdx",
                          "W4_Paranoia_Total",
                          "W4_Conspiracy_Total",
                          "W4_TotalMigrantAttitudes")])

# Convert 'W4_Education_categories_Num1' from factor to numeric
cor_df1$W4_Education_categories_Num1 <- as.numeric(as.character(cor_df1$W4_Education_categories_Num1))

# Convert 'W4_Income_2019' from labelled variable to numeric
cor_df1$W4_Income_2019_1 <- as.numeric(labels(cor_df1$W4_Income_2019)[cor_df1$W4_Income_2019])

# Convert 'W4_PTSDdx' from labelled variable to numeric
cor_df1$W4_PTSDdx <- as.numeric(as.character(cor_df1$W4_PTSDdx))

# Give df new lables
new_names1 <- c("Age (Years)", "Male", "Visible minority", "Education", "Income", "Political scale", "Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openness","Depression", "PTSD_dx", "Paranoia", "Conspiracy", "Migrant Attitudes")
names(cor_df1) <- new_names1

# Create correlation matrix
cor_matrix1 <- cor(cor_df1, use = "pairwise.complete.obs")
# Calculate p-values for correlations
p_values <- cor.mtest(cor_df1, conf.level = 0.95)$p

# Create a correlation plot with colors, highlighting significant correlations
corrplot(cor_matrix1, method = "color", type = "upper", tl.col = "black", tl.srt = 45, p.mat = p_values,sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, insig = 'label_sig', pch.col = 'grey20', order = "AOE",number.cex = 0.7, tl.cex = 0.7, mar = c(0,0,2,0))

# Create a correlation plot with colors, highlighting  correlations coefficents
corrplot(cor_matrix1, method = "color", type = "upper", tl.col = "black", tl.srt = 45,order = "AOE",number.cex = 0.7, tl.cex = 0.7,mar = c(0,0,2,0))$corrPos -> p1
text(p1$x, p1$y, round(p1$corr, 2), cex = 0.5)


```

###  Create models:

```{r}
### Linear models:

m1<-lm(df$W4_TotalMigrantAttitudes ~ df$W4_Age_year + df$W4_Gender_Binary + df$W4_Visible_minority + df$W4_Education_categories + df$W4_Employment_categories + df$W4_Income_2019_categories + df$W4_Political_scale)
summary(m1)

m2<-lm(df$W4_TotalMigrantAttitudes ~ df$W4_Extraversion_Total + df$W4_Agreeable_Total + df$W4_Conscientious_Total + df$W4_Neuroticism_Total + df$W4_Openness_Total)
summary(m2)

m3<-lm(df$W4_TotalMigrantAttitudes ~ df$W4_Dep_Total + df$W4_PTSDdx + df$W4_Paranoia_Total + df$W4_Conspiracy_Total)
summary(m3)

m4<-lm(df$W4_TotalMigrantAttitudes ~ df$W4_Age_year + df$W4_Gender_Binary + df$W4_Visible_minority + df$W4_Education_categories + df$W4_Employment_categories + df$W4_Income_2019_categories + df$W4_Political_scale + df$W4_Extraversion_Total + df$W4_Agreeable_Total + df$W4_Conscientious_Total + df$W4_Neuroticism_Total + df$W4_Openness_Total +df$W4_Dep_Total + df$W4_PTSDdx + df$W4_Paranoia_Total + df$W4_Conspiracy_Total )
summary(m4)

# Export the models

stargazer(m1,m2,m3,m4)
```

### Calculate VIF:

```{r}
# Calculate VIF
vif_values <- vif(m4)
write.csv(vif_values, "vif_values.csv")
```

### Assumption plots:

```{r}
plot(m4)
```
